Open-Set Domain Adaptation through Self-Supervision

### 2022-02-12 paper/main.tex
Step II reviewed

\subsection{Approach}
\label{sec:apporach}
The proposed method is split in two sequential stages. First, to avoid negative transfer during the domain alignment step, we want a model
that is able to separate the target dataset into $\mathcal{D}_t^{unk}$ and $\mathcal{D}_t^{knw}$. To do that, we leverage the
power of the rotation pre-text task to perform the separation.
Next, in the second stage, we can close the gap between the source domain and the target domain exploiting $\mathcal{D}_t^{knw}$
using the same self-supervised task. Furthermore, we leverage $\mathcal{D}_t^{unk}$ to learn the additional {\it unknown} class.

@@

\subsection{Step II: Domain Alignment}
\label{sec:domain_alignment}

@@

In this phase,
we build two "new" datasets.
The first one is $\mathcal{D}_s^*$, composed as $\mathcal{D}_s \cup \mathcal{D}_t^{unk}$.
Since $\mathcal{D}_t$ is an unlabeled dataset but $\mathcal{D}_s^*$ is a labeled one,
samples in $\mathcal{D}_t^{unk}$ are labeled as unknowns.
The second one is $\mathcal{D}_t^{knw}$.
Here we also use two new classifiers,
$C_2$ and $R_2$,
similarly to how they were used in the first phase but with $C_2$ predicting $|\mathcal{C}_s|+1$ classes to accomodate unknown predictions.
In this step $R_2$ is always single-headed and has its own weight $\alpha_2$.
The objective function is the same described in equations \ref{eq:baseloss_class} and \ref{eq:baseloss_rot}.
The main difference here is that $C_2$ is actually trained to recognize unknown classes,
on both domains.

>>>>

\subsection{Approach}
\label{sec:apporach}
The proposed method is split in two sequential stages. First, to avoid negative transfer during the domain alignment step, we want a model
that is able to separate the target dataset into $\mathcal{D}_t^{unk}$, which contains images belonging to unknown classes,
and $\mathcal{D}_t^{knw}$, which contains only images belonging to the known classes. To do that, we leverage the
power of the rotation pre-text task to perform the separation.
Next, in the second stage, we can close the gap between the source domain and the target domain exploiting $\mathcal{D}_t^{knw}$
using the same self-supervised task. Furthermore, we leverage $\mathcal{D}_t^{unk}$ to learn the additional {\it unknown} class.

@@

\subsection{Step II: domain alignment}
\label{sec:domain_alignment}

@@

In this step,
having separated the target into a known part $\mathcal{D}_t^{knw}$,
and an unknown part $\mathcal{D}_t^{unk}$,
we arrange two new datasets in order to perform the domain alignment while also learning
the unknown class.
The first one is $\mathcal{D}_s^*$, composed as $\mathcal{D}_s \cup \mathcal{D}_t^{unk}$,
which contains the original source images plus the target images identified as unknown classes.
We thus set the labels for $\mathcal{D}_t^{unk}$ as the class {\it unknown}.
The second one is $\mathcal{D}_t^{knw}$, which can be used to perform the domain alignment
without the risk of negative transfer.
While the feature extractor $E$ is inherited from the previous stage and leverages the previous training phase,
we also use two new classifiers,
$C_2$ and $R_2$.
They are similar to the previous classifiers but they have two important differences.
$C_2$ now has a $(|\mathcal{C}_s|+1)$-dimensional output to accomodate also the unknown class predictions and also
benefits from the previous learning,
while $R_2$ is always a single-head rotation classifier and starts the learning from scratch.
The training phase is the same as before with the difference that we do not have the center loss this time.
We also employ a different hyperparameter $\alpha_2$ to weigh the rotation classifier loss contribution.
The objective function is again $\mathcal{L}_2 = \mathcal{L}_{C_2} + \mathcal{L}_{R_2}$,
where the two contributions are identical to equations \ref{eq:baseloss_class} and \ref{eq:baseloss_rot}.
We report the $\mathcal{L}_{R_2}$ loss contributions to make clear the usage of $\alpha_2$ and we
recall that $\mathcal{L}_{C_2}$ is now computed using the new arranged dataset $\mathcal{D}_s^*$.

\begin{equation}
  \mathcal{L}_{R_2} = -\alpha_2\sum_{i\in\mathcal{D}_t^{knw}} {\bf z}_i^s \log \hat{{\bf z}}_i^s
  \label{eq:baseloss_rot2}
\end{equation}
